## Part one
a)

--Import the required libraries.
```{r}
library(ggplot2)
```

-- Define the target density function f(x).
```{r}
f <- function(x) {
  0.5 * exp(-abs(x))
}
```
-- Create variables N = 10000 and S = 1 as given in the question.Let x0 = 0 and the first term of sample_space to be x0 with n numbers of 0.

```{r}
N <- 10000
s <- 1
x0 <- 0
sample_space <- numeric(N)
sample_space[1] <- x0
```

-- Function for Metropolis-Hastings algorithm.
```{r}
for (i in 2:N) {
  x_current <- sample_space[i - 1]
  x_star <- rnorm(1, mean = x_current, sd = s)
  ratio <- log(f(x_star)) - log(f(x_current))
  if (runif(1) < exp(ratio)) {
    sample_space[i] <- x_star
  } else {
    sample_space[i] <- x_current
  }
}

samples_df <- data.frame(x = sample_space)
```


--Using the values generated, construct a histogram and kernel density plot and overlay graph of f(x).
```{r}
ggplot(samples_df, aes(x = x)) +
  geom_histogram(aes(y = after_stat(density)), bins = 50, fill = "skyblue", color = "black", alpha = 0.5) +
  geom_density(color = "blue", linewidth = 1.2) +
  stat_function(fun = f, color = "red", linewidth = 1.2) +
  labs(title = "Metropolis-Hastings Simulation", x = "x", y = "Density") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
```

--Using the functions defined above, sub in values of N and S to generate samples to find sample mean and standard deviation.
```{r}
sample_mean <- mean(sample_space)
sample_sd <- sd(sample_space)
cat("Sample Mean:", sample_mean, "\n")
cat("Sample Standard Deviation:", sample_sd, "\n")
```

## Part B


-- Function to run Metropolis-Hastings for chains.
```{r}
run_chain <- function(N, s, x0) {
  sample_space <- numeric(N)
  for (i in 2:N) {
    x_current <- sample_space[i - 1]
    x_star <- rnorm(1, mean = x_current, sd = s)
    acceptance_ratio <- f(x_star) / f(x_current)
    if (runif(1) < acceptance_ratio) {
      sample_space[i] <- x_star
    } else {
      sample_space[i]<- x_current
    }
  }
  return(sample_space)
}
```

--Change parameter N,J and s to new values given. Generate values for s in the interval between 0.001 and 1.
```{r}
N <- 2000
J <- 4
s_values <- seq(0.001, 1, length.out = 150)
R_hat_values <- numeric(length(s_values))
```

--Generate values for r hat and store it in array R_hat_values.
```{r}
for (k in seq_along(s_values)) {
  s <- s_values[k]
  chains <- replicate(J, run_chain(N, s, runif(1)), simplify = FALSE)
  M_j <- sapply(chains, mean)
  V_j <- sapply(chains, var)
  W <- mean(V_j)
  M <- mean(M_j)
  B <- sum((M_j - M)^2)
  R_hat <- sqrt((B + W) / W)
  R_hat_values[k] <- R_hat
}
```
--plotting the graph of values of rhat over a grid of s values in the interval between 0.001 and 1.
```{r}
R_hat_df <- data.frame(s = s_values, R_hat = R_hat_values)
ggplot(R_hat_df, aes(x = s, y = R_hat)) +
  geom_line(color = "blue") +
  geom_hline(yintercept = 1.05, color = "red", linetype = "dashed", linewidth = 1, alpha = 0.8) +
  labs(x = "s", y = expression(hat(R)),
       title = expression(paste("Convergence Diagnostic ", hat(R), " over different values of s"))) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

cat("R_hat Value:", R_hat, "\n")
```

