
# Question 2(a)

Import libraries
```{r}
library(dplyr)
library(ggplot2)
library(ggcorrplot)
library(caret)
library(pROC)
library(ROCR)
library(tidyr)

```

Reading Dataset
```{r}
# Create an empty list to store data frames
df_list <- list()

# Goes through each year from 1995 to 1999
for (year in 1995:1999) {
  file_path <- paste0("dataset/", year, ".csv")
  
  if (file.exists(file_path)) {
    df <- read.csv(file_path)

    # Data Cleaning
    df <- df[!duplicated(df), ]  # Remove duplicates

    # Remove negative values and Null values for time
    df <- df[df$ActualElapsedTime >= 0 | is.na(df$ActualElapsedTime), ]
    df <- df[df$CRSElapsedTime >= 0 | is.na(df$CRSElapsedTime), ]

    # Remove values greater than 2359 and Null values for 24 hour time
    df <- df[df$DepTime <= 2359 | is.na(df$DepTime), ]
    df <- df[df$ArrTime <= 2359 | is.na(df$ArrTime), ]

    # Append cleaned data to the main DataFrame
    df_list[[length(df_list) + 1]] <- df
  }
}

# Combine all yearly cleaned data into one data frame
main_df <- do.call(rbind, df_list)

# View first few rows
head(main_df)

```

```{r}
list.files("dataset")
```
Converting main_df column names to lowercase.
```{r}
colnames(main_df) <- tolower(colnames(main_df))

print(colnames(main_df))
```
# Creating delay_df to answer part (a)

Taking columns 'year', 'month', 'dayofmonth', 'dayofweek', 'deptime', 'arrtime', 'arrdelay' into delay_df
```{r}
# Select specific columns and create a new data frame
delay_df <- main_df[, c("year", "month", "dayofmonth", "dayofweek", "deptime", "arrtime", "arrdelay")]

# View the first few rows
head(delay_df)

```

Checking for null values
```{r}
# Count missing (NA) values for each column
colSums(is.na(delay_df))
```

Summary of delay_df
```{r}
str(delay_df)
```
Converting 'month' and 'dayofweek' into their names for easier identification
```{r}
# checking the values in 
unique(delay_df$month)
unique(delay_df$dayofweek)

```

```{r}
#creating a mapping dataframe for 'dayofweek'
day_of_week_mapping <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")

#Creating a mapping dataframe for 'month' 
month_mapping <- c("January", "February", "March", "April", "May", "June", "July", "August", "September", "October", "November", "December")

# Map numeric values to strings
delay_df$dayofweekstring <- day_of_week_mapping[delay_df$dayofweek]
delay_df$monthstring <- month_mapping[delay_df$month]

# View result
head(delay_df)
```
Check "monthstring" and "dayofweekstring" for null values
```{r}
sum(is.na(delay_df$monthstring))
sum(is.na(delay_df$dayofweek))
```

# Finding Delays

Based on United States Federal Aviation Administration (FAA), a flight is considered to be delayed when it's arrival time is 15 minutes later than its scheduled time. For part (a), I will be caterizing the delays by the following:

- on_time: less than 15mins delayed

- mild_delay: 15 to 60mins delayed

- moderate_delay: 60 to 180mins delayed 

- severe_delay: more than 180mins delayed 

These will be added to a new column "delay_classification" and will provide on insights that will help us find out what are the best times of the weeks to book an air ticket. 


```{r}
delay_classification <- function(time) {
  if (is.na(time)) {
    return(NA)#If the value is missing
  } else if (time <= 15) { #Delay is less than 15 minutes
    return("on_time")
  } else if (time < 60) { #Delay is less than 60 minutes
    return("mild_delay")
  } else if (time < 180) { #Delay is less than 180 minutes
    return("moderate_delay")
  } else {
    return("severe_delay") #Delay is greater than 180 minutes
  }
}

delay_df$delay_classification <- sapply(delay_df$arrdelay, delay_classification)
print(head(delay_df))
```
Removing Null values in delay_classification
```{r}
delay_df <- delay_df[!is.na(delay_df$delay_classification), ]
```

Checking for missing values in delay_classification
```{r}
sum(is.na(delay_df$delay_classification))
```
Checking delay_classification distribution
```{r}
table(delay_df$delay_classification)
```
# Classifying time of day

We will be grouping the time of day as following:

- Midnight: 0000 to 0559

- Morning: 0600 to 1159

- Afternoon: 1200 to 1759

- Night: 1800 to 2359 

These will be added to a new column "time_of_day" and will provide insights on when is the best time of the day to book an air ticket. We used deptime as the timing on tickets are usually the departure time of flights.
```{r}
# Creating a function for time of day classification
time_of_day <- function(time) {
  if (is.na(time)) {
    return(NA)
  } else if (time <= 559) {
    return("Midnight")
  } else if (time <= 1159) {
    return("Morning")
  } else if (time <= 1759) {
    return("Afternoon")
  } else {
    return("Night")
  }
}

delay_df$time_of_day <- sapply(delay_df$deptime, time_of_day)

print(head(delay_df))
```
Check for missing values in time_of_day
```{r}
sum(is.na(delay_df$time_of_day))
```
Replacing Month and Days with their respective names (string)
```{r}
# Creating a mapping list for days of a week and applying it to delay_df
days <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")
delay_df$dayofweekstring <- days[delay_df$dayofweek]

# Creating a mapping list for months
months <- c("January", "February", "March", "April", "May", "June", "July", "August", "September", "October", "November", "December")
delay_df$monthstring <- months[delay_df$month]

# Finalised main data frame
head(delay_df)

```
Checking for null values for before using dataframe to answer 2(a)
```{r}
colSums(is.na(delay_df))
```
Renaming dataframe to final_delay_df
```{r}
final_delay_df <- delay_df
```

# 2(a) Finding the best time of day to book a plane ticket

In order to find the best time of day to book a plane ticket, we will be grouping delays by time_of_day and 'delay_classification" and finding the count of each.
```{r}
# Count combinations of time_of_day and delay_classification
grouped_time_df <- as.data.frame(table(final_delay_df$time_of_day, final_delay_df$delay_classification))

# Rename columns for clarity
colnames(grouped_time_df) <- c("time_of_day", "delay_classification", "count")

# View result
print(grouped_time_df)
```

```{r}
# calculating the total count of delays for each time of day
grouped_time_df <- grouped_time_df %>%
  group_by(time_of_day) %>%
  mutate(
    total_time_of_day = sum(count), # total count of delays for each time of day
    percentage = round((count / total_time_of_day) * 100, 2) # compute percentage
  ) %>%
  ungroup()

# Display the result
print(grouped_time_df)

```

Plotting a pie chart that shows the proportion of each delay classification. 
```{r}
# Aggregate the count for each delay classification
delay_counts <- aggregate(count ~ delay_classification, data = grouped_time_df, sum)

# Define colors for delay classifications (from green to red)
color_map <- c(
  "on_time" = "green",        
  "mild_delay" = "yellow",     
  "moderate_delay" = "orange", 
  "severe_delay" = "red"    
)
# Match the order of labels to the color map
colors <- color_map[match(delay_counts$delay_classification, names(color_map))]

pie(
  delay_counts$count,
  labels = paste0(delay_counts$delay_classification, " (", round(100 * delay_counts$count / sum(delay_counts$count), 1), "%)"),
  col = colors,
  main = "Distribution of Delay Classification"
)


```
Plotting a bar chart that shows percentage of delays against time of day.
```{r}
# Optional: reorder delay_classification for consistent color assignment
grouped_time_df$delay_classification <- factor(
  grouped_time_df$delay_classification,
  levels = c("on_time", "mild_delay", "moderate_delay", "severe_delay")
)

# Define color mapping similar to Python
color_map <- c(
  "on_time" = "green",       
  "mild_delay" = "yellow",    
  "moderate_delay" = "orange",
  "severe_delay" = "red"  
)

# Plot
ggplot(grouped_time_df, aes(x = time_of_day, y = percentage, fill = delay_classification)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8)) +
  geom_text(aes(label = paste0(percentage, "%")),
            position = position_dodge(width = 0.8),
            vjust = -0.3, size = 3.5) +
  scale_fill_manual(values = color_map) +
  labs(title = "Percentage of Delays by Time of Day",
       x = "Time of Day", y = "Percentage",
       fill = "Delay Classification") +
  theme_minimal()
```
Based on the graph above, morning flights have the most flights that are on time at 86.68% of flights, while night flights tend to be delayed more often with only 70.69% of flights being on time. Based on the visuals above, I would recommend booking tickets in the mornings and avoiding booking night flights.


# Finding the best day of the week to book flight tickets.

 In order to find the best day of the week to book a plane ticket, we will be grouping delays by "dayofweekstring" and 'delay_classification" and finding the count of each.
```{r}
# Grouping by day of week and delay classification, and counting
grouped_day_df <- final_delay_df %>%
  group_by(dayofweekstring, delay_classification) %>%
  summarise(count = n(), .groups = 'drop')

# Calculate percentages within each day of the week
grouped_day_df <- grouped_day_df %>%
  group_by(dayofweekstring) %>%
  mutate(percentage = round((count / sum(count)) * 100, 2)) %>%
  ungroup()

# View the result
print(grouped_day_df)
```
Plotting a bar chart that shows percentage of delays against day of the week.
```{r}
final_delay_df$dayofweekstring <- factor(
  final_delay_df$dayofweekstring,
  levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")
)

# Group, count, and calculate percentages
grouped_day_df <- final_delay_df %>%
  group_by(dayofweekstring, delay_classification) %>%
  summarise(count = n(), .groups = 'drop') %>%
  group_by(dayofweekstring) %>%
  mutate(percentage = round((count / sum(count)) * 100, 2)) %>%
  ungroup()

# Define color map
color_map <- c(
  "on_time" = "green",        # Green
  "mild_delay" = "yellow",     # Yellow
  "moderate_delay" = "orange", # Orange
  "severe_delay" = "red"    # Red
)

# Create bar plot
ggplot(grouped_day_df, aes(x = dayofweekstring, y = percentage, fill = delay_classification)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label = paste0(percentage, "%")), 
            position = position_dodge(width = 0.9), 
            vjust = -0.2, size = 3) +
  scale_fill_manual(values = color_map) +
  labs(title = "Percentage of Delays by Day of the Week",
       x = "Day of the Week",
       y = "Percentage",
       fill = "Delay Classification") +
  theme_minimal()

```
Based on the graph above, Saturdays has the highest percentage of flights that are on time at 82.98% while Fridays has the least percentage of flights that are on time at 76.12%. 

Based on the visual above, I would recommend booking flights on Saturdays while avoiding booking flights on Fridays.

# Finding best days of the week to book a plane ticket for each year
```{r}
# Group by year and day of week, and calculate average arrival delay
grouped_days <- final_delay_df %>%
  group_by(year, dayofweekstring) %>%
  summarise(avg_arrdelay = mean(arrdelay, na.rm = TRUE), .groups = "drop")

# Find the best day (with minimum average delay) for each year
best_days <- grouped_days %>%
  group_by(year) %>%
  filter(avg_arrdelay == min(avg_arrdelay)) %>%
  ungroup()

# Print results
for (i in 1:nrow(best_days)) {
  year <- best_days$year[i]
  day <- best_days$dayofweekstring[i]
  delay <- round(best_days$avg_arrdelay[i], 2)
  cat(sprintf("In %d, the best day to fly was %s with an average arrival delay of %.2f minutes.\n", year, day, delay))
}

```
# Finding best time of day to book a plane ticket for each year
```{r}
# Group by year and time of day, calculate average arrival delay
grouped_time <- final_delay_df %>%
  group_by(year, time_of_day) %>%
  summarise(avg_arrdelay = mean(arrdelay, na.rm = TRUE), .groups = "drop")

# For each year, find the time of day with the lowest average delay
best_times <- grouped_time %>%
  group_by(year) %>%
  filter(avg_arrdelay == min(avg_arrdelay)) %>%
  ungroup()

# Print results for each year
for (i in 1:nrow(best_times)) {
  year <- best_times$year[i]
  time <- best_times$time_of_day[i]
  delay <- round(best_times$avg_arrdelay[i], 2)
  cat(sprintf("In %d, the best time of day to fly was %s with an average arrival delay of %.2f minutes.\n", year, time, delay))
}

```
# Part 2(b) Evaluate wether older planes suffer more delays on a year-to-year basis.

Loading dataset
```{r}
planes_data_df <- read.csv("dataset/plane-data.csv")
print(planes_data_df)
```

Checking for null values and data type
```{r}
colSums(is.na(planes_data_df))
str(planes_data_df)
planes_data_df
```

Converting tailnums in plane_data_df and main_df to lowercase 
```{r}
planes_data_df$tailnum <- tolower(planes_data_df$tailnum)
main_df$tailnum <- tolower(main_df$tailnum)
```

Combining main_df and planes_data_df using 'tailnum'.
```{r}
merged_df <- main_df %>%
  left_join(planes_data_df, by = "tailnum", suffix = c("main_df", "planes_data_df"))

merged_df
```
Checking for null values in merged_df
```{r}
colSums(is.na(merged_df))
```

```{r}
merged_df
```
We will be choosing and renaming some columns to answer the second question.

Same as part (a), we will be using arrival delay only as DOT defines delay as arrival delay.
```{r}
age_df <- merged_df[, c("yearmain_df", "flightnum", "tailnum", "arrdelay", "yearplanes_data_df")]
head(age_df)
```
Rename yearmain_df to current_year and yearplanes_data_df to plane_year
```{r}
colnames(age_df)[colnames(age_df) == "yearmain_df"] <- "current_year"
colnames(age_df)[colnames(age_df) == "yearplanes_data_df"] <- "plane_year"

head(age_df)
```
Removing null values from plane_year
```{r}
age_df <- age_df[!is.na(age_df$plane_year), ]

colSums(is.na(age_df))
```
To find the age of a plane, we will be taking current year minus the issue date.
```{r}
#converting plane_year to integers before calculation
age_df$plane_year <- as.integer(age_df$plane_year)

#calculating plane age using current year minus issue date 
age_df$plane_age <- age_df$current_year - age_df$plane_year
head(age_df)


```

```{r}
colSums(is.na(age_df))
```

Removing null values in arrival delays and plane age
```{r}
age_df <- age_df[!is.na(age_df$arrdelay) & !is.na(age_df$plane_age), ]
colSums(is.na(age_df))

```

# General Trend

We will now be grouping the dataset by the plane's tail number to get the average delays for each model of plane.
```{r}
plane_summary_df <- age_df %>%
  group_by(tailnum) %>%
  summarise(
    average_arrdelay = mean(arrdelay, na.rm = TRUE),
    plane_age = first(plane_age)                       
  )
head(plane_summary_df)
```
Removing plane ages that are less than 0
```{r}
plane_summary_df <- plane_summary_df %>%
  filter(plane_age >= 0)

head(plane_summary_df)
```
We will now be grouping the dataset by the plane's age. 
```{r}
plane_age_groups <- plane_summary_df %>%
  group_by(plane_age) %>%
  summarise(average_arrdelay = mean(average_arrdelay, na.rm = TRUE))
head(plane_age_groups)
```

Plotting average arrival delay against plane age to see the general trend. 
```{r}
ggplot(plane_age_groups, aes(x = plane_age, y = average_arrdelay)) +
  geom_line(color = "red") +
  geom_point(color = "red") +
  labs(
    title = "Average Arrival Delay by Plane Age",
    x = "Plane Age",
    y = "Average Arrival Delay"
  ) +
  theme_minimal()
```
Based on the graph above, from age 0 to 24, the general trend of delays against plane age is decreasing, meaning as age increases, delay decreases. However, after the age of 24, delay is generally increasing as age increases.


```{r}
ggplot(plane_age_groups, aes(x = plane_age, y = average_arrdelay)) +
  geom_point() +  # Scatterplot points
  geom_smooth(method = "lm", color = "red") +  # Linear regression line
  labs(
    title = "Plane Age vs Average Arrival Delay with Trend Line",
    x = "Plane Age",
    y = "Average Arrival Delay"
  ) +
  theme_minimal()

```

```{r}
# Compute Spearman correlation
corr_matrix <- cor(plane_age_groups, method = "spearman", use = "complete.obs")

# Plot
ggcorrplot(corr_matrix, lab = TRUE, lab_size = 4, 
           colors = c("blue", "white", "red"),
           title = "Feature Correlation Heatmap")

```
This shows that plane age has a negative correlation between arrival delays, which supports the graphs above, where as plane age increases, delays generally decrease. 

# Trend throughout the years.

We will first group the dataset by its years.
```{r}
# Create an empty list to store the yearly DataFrames
yearly_age_dfs <- list()

# Loop through the years and create a filtered DataFrame for each year
for (year in 1995:1999) {
  yearly_age_dfs[[as.character(year)]] <- age_df[age_df$current_year == year, ]
}

# Check the result for one of the years
head(yearly_age_dfs[["1995"]])
```
Grouping dataset by tail number
```{r}
# Create an empty list to store the yearly plane summaries
yearly_plane_summaries <- list()

# Loop through each year's DataFrame in the list
for (year in names(yearly_age_dfs)) {
  
  # Get the data frame for the current year
  df <- yearly_age_dfs[[year]]
  
  # Summarize the data by 'tailnum'
  summary <- df %>%
    group_by(tailnum) %>%
    summarise(
      average_arrdelay = mean(arrdelay, na.rm = TRUE),
      plane_age = first(plane_age)
    ) %>%
    ungroup()  # To remove the grouping
  
  # Store the summary in the list with the year as the key
  yearly_plane_summaries[[year]] <- summary
}

# Check the result for one of the years
head(yearly_plane_summaries[["1995"]])
```
Removing plane age that are less than 0
```{r}
# Loop through each year in the yearly_plane_summaries list
for (year in names(yearly_plane_summaries)) {
  
  # Filter the summary for the current year where plane_age >= 0
  yearly_plane_summaries[[year]] <- yearly_plane_summaries[[year]] %>%
    filter(plane_age >= 0)
}

# Check the result for one of the years
head(yearly_plane_summaries[["1995"]])

```
Grouping the dataset by its plane age
```{r}
# Dictionary to store grouped-by-age DataFrames
yearly_age_groups <- list()

# Loop through each year's summary DataFrame
for (year in names(yearly_plane_summaries)) {
  
  # Group by plane_age and calculate the mean of average_arrdelay
  grouped <- yearly_plane_summaries[[year]] %>%
    group_by(plane_age) %>%
    summarise(average_arrdelay = mean(average_arrdelay, na.rm = TRUE)) %>%
    ungroup()
  
  # Store the grouped data in the list
  yearly_age_groups[[year]] <- grouped
}

# Check the result for one of the years
head(yearly_age_groups[["1995"]])

```

```{r}
# Loop through each year's grouped data
for (year in names(yearly_age_groups)) {
  
  # Create the plot using ggplot2
  p <- ggplot(yearly_age_groups[[year]], aes(x = plane_age, y = average_arrdelay)) +
    geom_point(size = 2) +  # Scatter plot
    geom_smooth(method = "lm", color = "red", se = FALSE) +  # Add regression line
    ggtitle(paste("Plane Age vs Average Arrival Delay (", year, ")", sep = "")) +
    xlab("Plane Age (Years)") +
    ylab("Average Arrival Delay (Minutes)") +
    theme_minimal()
  
  # Print the plot
  print(p)
}
```


Plot Correlation Matrix
```{r}
# Loop through each year's grouped data
for (year in names(yearly_age_groups)) {
  
  # Extract the relevant dataframe for the current year
  df <- yearly_age_groups[[year]]
  
  # Compute Spearman correlation matrix (for the two columns: plane_age and average_arrdelay)
  corr_matrix <- cor(df, method = "spearman", use = "complete.obs")

  # Plot the correlation matrix using ggcorrplot
  corr_plot <- ggcorrplot(corr_matrix, 
                          lab = TRUE,             # Display correlation coefficients on the heatmap
                          lab_size = 4,           # Size of correlation values
                          colors = c("blue", "white", "red"),  # Color gradient
                          title = paste("Feature Correlation Heatmap (", year, ")", sep = ""), 
                          method = "square")      # Square-shaped cells
  
  # Display the plot
  print(corr_plot)
}

```
These heatmaps support the earlier visual observations that the trend between plane age and average arrival delay is generally downward. In 1995, 1996, and 1997, the Spearman correlation between plane age and average arrival delay ranges from -0.44 to -0.56, indicating a moderate negative correlation. This suggests that older aircraft were associated with lower average arrival delays during these years. In contrast, 1998 shows a stronger negative correlation of -0.67, while 1999 shows a weaker negative correlation of -0.35, indicating that the relationship varied slightly from year to year but remained consistently negative across the five-year period.


Reasons why this might be the case:

- newer planes being assigned to longer and more complicated routes while older planes being assigned to shorter and simpler routes, leading to newer planes having more delays compared to older planes

- older planes may be used much lesser than newer planes, leading to more instances of delays for newer planes

- there may be a survivorship bias as the older planes still flying are more reliable than the newer ones. 


# Part 2(c) Logistic Regression for the probability of diverted US flights

Data Cleaning 

We will be first be removing cancelled flights as cancelled flights can't be diverted.
```{r}
diverted_df <- main_df
diverted_df <- diverted_df[diverted_df$cancelled == 0, ]
table(diverted_df$cancelled)
```
Removing null values in columns that we will be using for logistic regression.
```{r}
colSums(is.na(diverted_df))

```

```{r}
features <- c('month', 'dayofmonth', 'dayofweek', 'crsdeptime', 'crsarrtime', 'distance','uniquecarrier', 'origin', 'dest', 'depdelay', 'taxiout')

target <- 'diverted'

diverted_df <- diverted_df[!is.na(diverted_df[[target]]) & !apply(diverted_df[features], 1, anyNA), ]

colSums(is.na(diverted_df))
```
We will be first checking for perfect multicollinearity to identify and remove redundant variables.
```{r}
numerical_variables <- c('month', 'dayofmonth', 'dayofweek', 'crsdeptime', 'crsarrtime', 'distance', 'depdelay', 'taxiout')

colSums(is.na(diverted_df[numerical_variables]))
```

```{r}
library(ggcorrplot)

# Assign numerical variables to corr_df
corr_df <- diverted_df[, numerical_variables]

# Compute the Pearson correlation matrix
corr_matrix <- cor(corr_df, method = "pearson", use = "complete.obs")

# Plot the correlation heatmap
ggcorrplot(corr_matrix, lab = TRUE, lab_size = 4, 
           colors = c("blue", "white", "red"),
           title = "Pearson Correlation Heatmap (Numeric Features)")

```
Since there is no pairwise correlation that is equal to one, the assumption of no perfect multicollinerity is satisfied. However, we will be removing dayofmonth as it has very low correlation and has low predictive value.

# Logistic Regression Analysis throughout the years

For R, we will be splitting the dataset into its years so that there is enough memory to run the code. Firstly, we will maximise memory by clearing all the other objects before running the logistic regression for the year. The coefficient will then be downloaded as a csv before repeating the steps again for the next year. At the end, we will then add all the coefficients into a dataframe for visualisations

```{r}
diverted_df
```
Selecting the columns 
```{r}
# Using features and target that are defined earlier
model_df <- diverted_df %>%
  select(all_of(features), all_of(target),year)
model_df
```


```{r}
print(colSums(is.na(model_df)))
```
Splitting the dataset by its year
```{r}
five_df <- model_df %>% filter(year == 1995)
six_df <- model_df %>% filter(year == 1996)
seven_df <- model_df %>% filter(year == 1997)
eight_df <- model_df %>% filter(year == 1998)
nine_df <- model_df %>% filter(year == 1999)

head(five_df)

```

```{r}
# Save the DataFrame as a CSV file
#write.csv(five_df, file = "dataset/five_df.csv", row.names = FALSE)
#write.csv(six_df, file = "dataset/six_df.csv", row.names = FALSE)
#write.csv(seven_df, file = "dataset/seven_df.csv", row.names = FALSE)
#write.csv(eight_df, file = "dataset/eight_df.csv", row.names = FALSE)
#write.csv(nine_df, file = "dataset/nine_df.csv", row.names = FALSE)

```


Maximising the amount of memory by removing other variables that are not needed 
```{r}
# List of all objects currently in the environment
all_objects <- ls()

# keep 1995 dataset
items_to_keep <- c("five_df")

# Identify objects to remove by subtracting the ones you want to keep
objects_to_remove <- setdiff(all_objects, items_to_keep)

# Remove the unnecessary objects from the environment
rm(list = objects_to_remove)

# Check that only the necessary objects remain
print(ls())

```

1995 Logistic Regression
```{r}
# Define categorical and numerical features
cat_features <- c('month', 'dayofweek', 'uniquecarrier', 'origin', 'dest')
num_features <- c('crsdeptime', 'crsarrtime', 'distance', 'depdelay', 'taxiout')
target <- 'diverted'
```

```{r}
# Define categorical and numerical features
cat_features <- c('month', 'dayofweek', 'uniquecarrier', 'origin', 'dest')
num_features <- c('crsdeptime', 'crsarrtime', 'distance', 'depdelay', 'taxiout')
target <- 'diverted'

# Prepare the dataset for the year 1995 (five_df)
df_year <- five_df %>%
  drop_na(all_of(c(cat_features, num_features, target)))

# Check if there's data after filtering
if (nrow(df_year) == 0) {
  stop("No data for year 1995")
}

# Split into X (features) and y (target)
X <- df_year[, c(cat_features, num_features)]
y <- df_year[, target]

# Stratified split with 30% of data
set.seed(42)
train_index <- createDataPartition(y, p = 0.3, list = FALSE, times = 1)  # 30% split with stratification
X_train <- X[train_index, ]
y_train <- y[train_index]
X_test <- X[-train_index, ]
y_test <- y[-train_index]

# Preprocessing: Center and scale numeric features, and apply one-hot encoding to categorical features
preProc <- preProcess(X_train, method = c("center", "scale"))
X_train <- predict(preProc, X_train)
X_test <- predict(preProc, X_test)

# Apply one-hot encoding for categorical features
X_train_cat <- model.matrix(~ month + dayofweek + uniquecarrier + origin + dest - 1, data = X_train)
X_test_cat <- model.matrix(~ month + dayofweek + uniquecarrier + origin + dest - 1, data = X_test)

# Combine the numerical and encoded categorical features
X_train_final <- cbind(X_train[, num_features], X_train_cat)
X_test_final <- cbind(X_test[, num_features], X_test_cat)

# Logistic Regression Model
model <- glm(y_train ~ ., data = as.data.frame(X_train_final), family = binomial(link = 'logit'))

# Predictions and probabilities
y_pred <- predict(model, as.data.frame(X_test_final), type = "response")
y_pred_class <- ifelse(y_pred > 0.5, 1, 0)

# Evaluate the model using confusion matrix, accuracy, and F1 score
confusion <- confusionMatrix(as.factor(y_pred_class), as.factor(y_test))

print("Year 1995 Results")
print(confusion)

# ROC Curve
roc_obj <- roc(y_test, y_pred)
plot(roc_obj, col = "blue", main = "ROC Curve for Year 1995")
print(paste("AUC:", auc(roc_obj)))

# Save coefficients for future analysis
coef_values <- coef(model)
coef_df <- data.frame(
  feature = names(coef_values),
  coefficient = coef_values
)

print(coef_df)


```
downloading coef_df
```{r}
write.csv(coef_df, file = "dataset/coef_df.csv", row.names = FALSE)
```

# Repeat the steps for 1996,1997,1998 and 1999

# 1996 Logistic Regression
Reading dataset 
```{r}
six_df <- read.csv("dataset/six_df.csv")
head(six_df)
```
Deleting objects that are not in use 
```{r}
# List of all objects currently in the environment
all_objects <- ls()

# keep 1995 dataset
items_to_keep <- c("six_df")

# Identify objects to remove by subtracting the ones you want to keep
objects_to_remove <- setdiff(all_objects, items_to_keep)

# Remove the unnecessary objects from the environment
rm(list = objects_to_remove)

# Check that only the necessary objects remain
print(ls())
```
1996 Logistic Regression 
```{r}
# Define categorical and numerical features
cat_features <- c('month', 'dayofweek', 'uniquecarrier', 'origin', 'dest')
num_features <- c('crsdeptime', 'crsarrtime', 'distance', 'depdelay', 'taxiout')
target <- 'diverted'

# Prepare the dataset for the year 1996 (six_df)
df_year <- six_df %>%
  drop_na(all_of(c(cat_features, num_features, target)))

# Check if there's data after filtering
if (nrow(df_year) == 0) {
  stop("No data for year 1996")
}

# Split into X (features) and y (target)
X <- df_year[, c(cat_features, num_features)]
y <- df_year[, target]

# Stratified split with 20% of data (without the 'stratify' argument, it works by default for factors)
set.seed(42)
train_index <- createDataPartition(y, p = 0.3, list = FALSE, times = 1)  # 30% split with stratification
X_train <- X[train_index, ]
y_train <- y[train_index]
X_test <- X[-train_index, ]
y_test <- y[-train_index]

# Preprocessing: Center and scale numeric features, and apply one-hot encoding to categorical features
preProc <- preProcess(X_train, method = c("center", "scale"))
X_train <- predict(preProc, X_train)
X_test <- predict(preProc, X_test)

# Apply one-hot encoding for categorical features
X_train_cat <- model.matrix(~ month + dayofweek + uniquecarrier + origin + dest - 1, data = X_train)
X_test_cat <- model.matrix(~ month + dayofweek + uniquecarrier + origin + dest - 1, data = X_test)

# Combine the numerical and encoded categorical features
X_train_final <- cbind(X_train[, num_features], X_train_cat)
X_test_final <- cbind(X_test[, num_features], X_test_cat)

# Logistic Regression Model
model <- glm(y_train ~ ., data = as.data.frame(X_train_final), family = binomial(link = 'logit'))

# Predictions and probabilities
y_pred <- predict(model, as.data.frame(X_test_final), type = "response")
y_pred_class <- ifelse(y_pred > 0.5, 1, 0)

# Evaluate the model using confusion matrix, accuracy, and F1 score
confusion <- confusionMatrix(as.factor(y_pred_class), as.factor(y_test))

print("Year 1996 Results")
print(confusion)

# ROC Curve
roc_obj <- roc(y_test, y_pred)
plot(roc_obj, col = "blue", main = "ROC Curve for Year 1996")
print(paste("AUC:", auc(roc_obj)))

# Save coefficients for future analysis
coef_values <- coef(model)
six_coef_df <- data.frame(
  feature = names(coef_values),
  coefficient = coef_values
)

print(six_coef_df)
```
downloading 'six_coef_df'
```{r}
write.csv(six_coef_df, file = "dataset/six_coef_df.csv", row.names = FALSE)
```

# 1997 Logistic Regression
Reading dataset 
```{r}
seven_df <- read.csv("dataset/seven_df.csv")
head(seven_df)
```

Deleting objects that are not in use 
```{r}
# List of all objects currently in the environment
all_objects <- ls()

# keep 1997 dataset
items_to_keep <- c("seven_df")

# Identify objects to remove by subtracting the ones you want to keep
objects_to_remove <- setdiff(all_objects, items_to_keep)

# Remove the unnecessary objects from the environment
rm(list = objects_to_remove)

# Check that only the necessary objects remain
print(ls())
```
1997 Logistic Regression 
```{r}
# Define categorical and numerical features
cat_features <- c('month', 'dayofweek', 'uniquecarrier', 'origin', 'dest')
num_features <- c('crsdeptime', 'crsarrtime', 'distance', 'depdelay', 'taxiout')
target <- 'diverted'

# Prepare the dataset for the year 1997 (seven_df)
df_year <- seven_df %>%
  drop_na(all_of(c(cat_features, num_features, target)))

# Check if there's data after filtering
if (nrow(df_year) == 0) {
  stop("No data for year 1997")
}

# Split into X (features) and y (target)
X <- df_year[, c(cat_features, num_features)]
y <- df_year[, target]

# Stratified split with 30% of data
set.seed(42)
train_index <- createDataPartition(y, p = 0.3, list = FALSE, times = 1)  # 30% split with stratification
X_train <- X[train_index, ]
y_train <- y[train_index]
X_test <- X[-train_index, ]
y_test <- y[-train_index]

# Preprocessing: Center and scale numeric features, and apply one-hot encoding to categorical features
preProc <- preProcess(X_train, method = c("center", "scale"))
X_train <- predict(preProc, X_train)
X_test <- predict(preProc, X_test)

# Apply one-hot encoding for categorical features
X_train_cat <- model.matrix(~ month + dayofweek + uniquecarrier + origin + dest - 1, data = X_train)
X_test_cat <- model.matrix(~ month + dayofweek + uniquecarrier + origin + dest - 1, data = X_test)

# Combine the numerical and encoded categorical features
X_train_final <- cbind(X_train[, num_features], X_train_cat)
X_test_final <- cbind(X_test[, num_features], X_test_cat)

# Logistic Regression Model
model <- glm(y_train ~ ., data = as.data.frame(X_train_final), family = binomial(link = 'logit'))

# Predictions and probabilities
y_pred <- predict(model, as.data.frame(X_test_final), type = "response")
y_pred_class <- ifelse(y_pred > 0.5, 1, 0)

# Evaluate the model using confusion matrix, accuracy, and F1 score
confusion <- confusionMatrix(as.factor(y_pred_class), as.factor(y_test))

print("Year 1997 Results")
print(confusion)

# ROC Curve
roc_obj <- roc(y_test, y_pred)
plot(roc_obj, col = "blue", main = "ROC Curve for Year 1997")
print(paste("AUC:", auc(roc_obj)))

# Save coefficients for future analysis
coef_values <- coef(model)
seven_coef_df <- data.frame(
  feature = names(coef_values),
  coefficient = coef_values
)

print(seven_coef_df)
```

downloading 'seven_coef_df'
```{r}
write.csv(seven_coef_df, file = "dataset/seven_coef_df.csv", row.names = FALSE)
head(seven_coef_df)
```

# 1998 Logistic Regression
Reading dataset 
```{r}
eight_df <- read.csv("dataset/eight_df.csv")
head(eight_df)
```
Deleting objects that are not in use 
```{r}
# List of all objects currently in the environment
all_objects <- ls()

# keep 1998 dataset
items_to_keep <- c("eight_df")

# Identify objects to remove by subtracting the ones you want to keep
objects_to_remove <- setdiff(all_objects, items_to_keep)

# Remove the unnecessary objects from the environment
rm(list = objects_to_remove)

# Check that only the necessary objects remain
print(ls())
```

1998 Logistic Regression 
```{r}
# Define categorical and numerical features
cat_features <- c('month', 'dayofweek', 'uniquecarrier', 'origin', 'dest')
num_features <- c('crsdeptime', 'crsarrtime', 'distance', 'depdelay', 'taxiout')
target <- 'diverted'

# Prepare the dataset for the year 1998 (eight_df)
df_year <- eight_df %>%
  drop_na(all_of(c(cat_features, num_features, target)))

# Check if there's data after filtering
if (nrow(df_year) == 0) {
  stop("No data for year 1998")
}

# Split into X (features) and y (target)
X <- df_year[, c(cat_features, num_features)]
y <- df_year[, target]

# Stratified split with 30% of data 
set.seed(42)
train_index <- createDataPartition(y, p = 0.3, list = FALSE, times = 1)  # 30% split with stratification
X_train <- X[train_index, ]
y_train <- y[train_index]
X_test <- X[-train_index, ]
y_test <- y[-train_index]

# Preprocessing: Center and scale numeric features, and apply one-hot encoding to categorical features
preProc <- preProcess(X_train, method = c("center", "scale"))
X_train <- predict(preProc, X_train)
X_test <- predict(preProc, X_test)

# Apply one-hot encoding for categorical features
X_train_cat <- model.matrix(~ month + dayofweek + uniquecarrier + origin + dest - 1, data = X_train)
X_test_cat <- model.matrix(~ month + dayofweek + uniquecarrier + origin + dest - 1, data = X_test)

# Combine the numerical and encoded categorical features
X_train_final <- cbind(X_train[, num_features], X_train_cat)
X_test_final <- cbind(X_test[, num_features], X_test_cat)

# Logistic Regression Model
model <- glm(y_train ~ ., data = as.data.frame(X_train_final), family = binomial(link = 'logit'))

# Predictions and probabilities
y_pred <- predict(model, as.data.frame(X_test_final), type = "response")
y_pred_class <- ifelse(y_pred > 0.5, 1, 0)

# Evaluate the model using confusion matrix, accuracy, and F1 score
confusion <- confusionMatrix(as.factor(y_pred_class), as.factor(y_test))

print("Year 1998 Results")
print(confusion)

# ROC Curve
roc_obj <- roc(y_test, y_pred)
plot(roc_obj, col = "blue", main = "ROC Curve for Year 1998")
print(paste("AUC:", auc(roc_obj)))

# Save coefficients for future analysis
coef_values <- coef(model)
eight_coef_df <- data.frame(
  feature = names(coef_values),
  coefficient = coef_values
)

print(eight_coef_df)
```

downloading 'eight_coef_df'
```{r}
write.csv(eight_coef_df, file = "dataset/eight_coef_df.csv", row.names = FALSE)
```

# 1999 Logistic Regression
Reading dataset 
```{r}
nine_df <- read.csv("dataset/nine_df.csv")
head(nine_df)
```
Deleting objects that are not in use 
```{r}
# List of all objects currently in the environment
all_objects <- ls()

# keep 1999 dataset
items_to_keep <- c("nine_df")

# Identify objects to remove by subtracting the ones you want to keep
objects_to_remove <- setdiff(all_objects, items_to_keep)

# Remove the unnecessary objects from the environment
rm(list = objects_to_remove)

# Check that only the necessary objects remain
print(ls())
```

1999 Logistic Regression 
```{r}
# Define categorical and numerical features
cat_features <- c('month', 'dayofweek', 'uniquecarrier', 'origin', 'dest')
num_features <- c('crsdeptime', 'crsarrtime', 'distance', 'depdelay', 'taxiout')
target <- 'diverted'

# Prepare the dataset for the year 1999 (nine_df)
df_year <- nine_df %>%
  drop_na(all_of(c(cat_features, num_features, target)))

# Check if there's data after filtering
if (nrow(df_year) == 0) {
  stop("No data for year 1999")
}

# Split into X (features) and y (target)
X <- df_year[, c(cat_features, num_features)]
y <- df_year[, target]

# Stratified split with 30% of data 
set.seed(42)
train_index <- createDataPartition(y, p = 0.3, list = FALSE, times = 1)  # 30% split with stratification
X_train <- X[train_index, ]
y_train <- y[train_index]
X_test <- X[-train_index, ]
y_test <- y[-train_index]

# Preprocessing: Center and scale numeric features, and apply one-hot encoding to categorical features
preProc <- preProcess(X_train, method = c("center", "scale"))
X_train <- predict(preProc, X_train)
X_test <- predict(preProc, X_test)

# Apply one-hot encoding for categorical features
X_train_cat <- model.matrix(~ month + dayofweek + uniquecarrier + origin + dest - 1, data = X_train)
X_test_cat <- model.matrix(~ month + dayofweek + uniquecarrier + origin + dest - 1, data = X_test)

# Combine the numerical and encoded categorical features
X_train_final <- cbind(X_train[, num_features], X_train_cat)
X_test_final <- cbind(X_test[, num_features], X_test_cat)

# Logistic Regression Model
model <- glm(y_train ~ ., data = as.data.frame(X_train_final), family = binomial(link = 'logit'))

# Predictions and probabilities
y_pred <- predict(model, as.data.frame(X_test_final), type = "response")
y_pred_class <- ifelse(y_pred > 0.5, 1, 0)

# Evaluate the model using confusion matrix, accuracy, and F1 score
confusion <- confusionMatrix(as.factor(y_pred_class), as.factor(y_test))

print("Year 1999 Results")
print(confusion)

# ROC Curve
roc_obj <- roc(y_test, y_pred)
plot(roc_obj, col = "blue", main = "ROC Curve for Year 1999")
print(paste("AUC:", auc(roc_obj)))

# Save coefficients for future analysis
coef_values <- coef(model)
nine_coef_df <- data.frame(
  feature = names(coef_values),
  coefficient = coef_values
)

print(nine_coef_df)
```

downloading 'nine_coef_df'
```{r}
write.csv(eight_coef_df, file = "dataset/eight_coef_df.csv", row.names = FALSE)
head(nine_coef_df)
```

Importing all the five coefficients from the years
```{r}
five_coef_df <- read.csv("dataset/coef_df.csv")
six_coef_df <- read.csv("dataset/six_coef_df.csv")
seven_coef_df <- read.csv("dataset/seven_coef_df.csv")
eight_coef_df <- read.csv("dataset/eight_coef_df.csv")
nine_coef_df <- read.csv("dataset/nine_coef_df.csv")
```

```{r}
six_coef_df
```
Adding a year column to coefficients dataframe
```{r}
five_coef_df$year <- 1995
six_coef_df$year <- 1996
seven_coef_df$year <- 1997
eight_coef_df$year <- 1998
nine_coef_df$year <- 1999
```

Adding all the coefficients into all_coef_df
```{r}
all_coef_df <- rbind(five_coef_df, six_coef_df, seven_coef_df, eight_coef_df, nine_coef_df)
```

Taking out 'distance', 'month','dayofweek','depdelay',taxiout','originAMA', 'uniquecarrierUA', 'destTVC', 'crsdeptime' and 'crsarrtime' 
```{r}
selected_vars <- c("distance", "month", "dayofweek", "depdelay", "taxiout",
                   "originAMA", "uniquecarrierUA", "destTVC", "crsdeptime", "crsarrtime")

final_coefficients_df <- all_coef_df[all_coef_df$feature %in% selected_vars, ]

```

Plot the coefficients over time 
```{r}
ggplot(final_coefficients_df, aes(x = year, y = coefficient, color = feature)) +
  geom_line() +
  geom_point() +
  labs(title = "Logistic Regression Coefficients Over Time",
       x = "Year",
       y = "Coefficient Value") +
  theme_minimal()

```

From the graphs and values above, we can see how the influence of different flight attributes on diversion probability evolves year to year. Features like distance and depdelay have consistent, strong positive effects, while others like month_2 and dayofweek_2 vary more in strength and direction.

```{r}
# Count the occurrences of diverted = 0 and 1
diverted_counts <- diverted_df %>%
  count(Diverted)
ggplot(diverted_counts, aes(x = as.factor(Diverted), y = n)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  geom_text(aes(label = n), vjust = -0.5) +
  labs(title = "Class Distribution: Diverted vs Not Diverted",
       x = "Diverted",
       y = "Count") +
  scale_y_continuous(labels = scales::comma) +
  theme_minimal()
```


As we can see, there is very little diverted compared to non-diverted flights in our dataset, which is causing our precision of our model to be extremely low for predicting diverted flights.  

Hence, the model can be improved using resampling techniques for imbalanced data, such as SMOTE or random oversampling. SMOTE, which stands for Synthetic Minority Over-sampling Technique generates fake examples by interpolating between existing minority class samples. Random Oversampling on the other hand duplicates examples from the minority class. Both these examples help to address the class imbalance and thus, would obtain a better precision score. 