
# Question 2(a)

Import libraries
```{r}
library(dplyr)
library(ggplot2)
library(ggcorrplot)
library(caret)
library(pROC)
library(ROCR)
library(tidyr)

```

Reading Dataset
```{r}
# Create an empty list to store data frames
df_list <- list()

# Goes through each year from 1995 to 1999
for (year in 1995:1999) {
  file_path <- paste0("dataset/", year, ".csv")
  
  if (file.exists(file_path)) {
    df <- read.csv(file_path)

    # Data Cleaning
    df <- df[!duplicated(df), ]  # Remove duplicates

    # Remove negative values and Null values for time
    df <- df[df$ActualElapsedTime >= 0 | is.na(df$ActualElapsedTime), ]

    # Remove negative values and Null values for time
    df <- df[df$CRSElapsedTime >= 0 | is.na(df$CRSElapsedTime), ]

    # #Remove values greater than 2359 and Null values for 24 hour time
    df <- df[df$DepTime <= 2359 | is.na(df$DepTime), ]
    df <- df[df$ArrTime <= 2359 | is.na(df$ArrTime), ]

    # Append cleaned data to the main DataFrame
    df_list[[length(df_list) + 1]] <- df
  }
}

# Display the first few rows of the final DataFrame
main_df <- do.call(rbind, df_list)

# View first few rows
head(main_df)
```

Converting main_df column names to lowercase.
```{r}
colnames(main_df) <- tolower(colnames(main_df))

print(colnames(main_df))
```
# Creating delay_df to answer part (a)

Taking columns 'year', 'month', 'dayofmonth', 'dayofweek', 'deptime', 'arrtime', 'arrdelay' into delay_df
```{r}
# Select specific columns and create a new data frame
delay_df <- main_df[, c("year", "month", "dayofmonth", "dayofweek", "deptime", "arrtime", "arrdelay")]

# View the first few rows
head(delay_df)

```

Checking for null values
```{r}
# Count missing (NA) values for each column
colSums(is.na(delay_df))

```

Summary of delay_df
```{r}
str(delay_df)
```
Converting 'month' and 'dayofweek' into their names for easier identification
```{r}
# checking the values in 
unique(delay_df$month)
unique(delay_df$dayofweek)

```

```{r}
#creating a mapping dataframe for 'dayofweek'
day_of_week_mapping <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")

#Creating a mapping dataframe for 'month' 
month_mapping <- c("January", "February", "March", "April", "May", "June", "July", "August", "September", "October", "November", "December")

# Map numeric values to strings
delay_df$dayofweekstring <- day_of_week_mapping[delay_df$dayofweek]
delay_df$monthstring <- month_mapping[delay_df$month]

# View result
head(delay_df)
```
Check "monthstring" and "dayofweekstring" for null values
```{r}
sum(is.na(delay_df$monthstring))
sum(is.na(delay_df$dayofweek))
```

# Finding Delays

Based on United States Federal Aviation Administration (FAA), a flight is considered to be delayed when it's arrival time is 15 minutes later than its scheduled time. For part (a), I will be caterizing the delays by the following:

- on_time: less than 15mins delayed

- mild_delay: 15 to 60mins delayed

- moderate_delay: 60 to 180mins delayed 

- severe_delay: more than 180mins delayed 

These will be added to a new column "delay_classification" and will provide on insights that will help us find out what are the best times of the weeks to book an air ticket. 


```{r}
delay_classification <- function(time) {
  if (is.na(time)) {
    return(NA)#If the value is missing
  } else if (time <= 15) { #Delay is less than 15 minutes
    return("on_time")
  } else if (time < 60) { #Delay is less than 60 minutes
    return("mild_delay")
  } else if (time < 180) { #Delay is less than 180 minutes
    return("moderate_delay")
  } else {
    return("severe_delay") #Delay is greater than 180 minutes
  }
}

delay_df$delay_classification <- sapply(delay_df$arrdelay, delay_classification)
print(head(delay_df))
```
Removing Null values in delay_classification
```{r}
delay_df <- delay_df[!is.na(delay_df$delay_classification), ]
```

Checking for missing values in delay_classification
```{r}
sum(is.na(delay_df$delay_classification))
```
Checking delay_classification distribution
```{r}
table(delay_df$delay_classification)
```
# Classifying time of day

We will be grouping the time of day as following:

- Midnight: 0000 to 0559

- Morning: 0600 to 1159

- Afternoon: 1200 to 1759

- Night: 1800 to 2359 

These will be added to a new column "time_of_day" and will provide insights on when is the best time of the day to book an air ticket. We used deptime as the timing on tickets are usually the departure time of flights.
```{r}
# Creating a function for time of day classification
time_of_day <- function(time) {
  if (is.na(time)) {
    return(NA)
  } else if (time <= 559) {
    return("Midnight")
  } else if (time <= 1159) {
    return("Morning")
  } else if (time <= 1759) {
    return("Afternoon")
  } else {
    return("Night")
  }
}

delay_df$time_of_day <- sapply(delay_df$deptime, time_of_day)

print(head(delay_df))
```
Check for missing values in time_of_day
```{r}
sum(is.na(delay_df$time_of_day))
```
Replacing Month and Days with their respective names (string)
```{r}
# Creating a mapping list for days of a week and applying it to delay_df
days <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")
delay_df$dayofweekstring <- days[delay_df$dayofweek]

# Creating a mapping list for months
months <- c("January", "February", "March", "April", "May", "June", "July", "August", "September", "October", "November", "December")
delay_df$monthstring <- months[delay_df$month]

# Finalised main data frame
head(delay_df)

```
Checking for null values for before using dataframe to answer 2(a)
```{r}
colSums(is.na(delay_df))
```
Renaming dataframe to final_delay_df
```{r}
final_delay_df <- delay_df
```

# 2(a) Finding the best time of day to book a plane ticket

In order to find the best time of day to book a plane ticket, we will be grouping delays by time_of_day and 'delay_classification" and finding the count of each.
```{r}
# Count combinations of time_of_day and delay_classification
grouped_time_df <- as.data.frame(table(final_delay_df$time_of_day, final_delay_df$delay_classification))

# Rename columns for clarity
colnames(grouped_time_df) <- c("time_of_day", "delay_classification", "count")

# View result
print(grouped_time_df)
```

```{r}
# calculating the total count of delays for each time of day
grouped_time_df <- grouped_time_df %>%
  group_by(time_of_day) %>%
  mutate(
    total_time_of_day = sum(count), # total count of delays for each time of day
    percentage = round((count / total_time_of_day) * 100, 2) # compute percentage
  ) %>%
  ungroup()

# Display the result
print(grouped_time_df)

```

Plotting a pie chart that shows the proportion of each delay classification. 
```{r}
# Aggregate the count for each delay classification
delay_counts <- aggregate(count ~ delay_classification, data = grouped_time_df, sum)

# Define colors for delay classifications (from green to red)
color_map <- c(
  "on_time" = "green",        
  "mild_delay" = "yellow",     
  "moderate_delay" = "orange", 
  "severe_delay" = "red"    
)
# Match the order of labels to the color map
colors <- color_map[match(delay_counts$delay_classification, names(color_map))]

pie(
  delay_counts$count,
  labels = paste0(delay_counts$delay_classification, " (", round(100 * delay_counts$count / sum(delay_counts$count), 1), "%)"),
  col = colors,
  main = "Distribution of Delay Classification"
)


```
Plotting a bar chart that shows percentage of delays against time of day.
```{r}
# Optional: reorder delay_classification for consistent color assignment
grouped_time_df$delay_classification <- factor(
  grouped_time_df$delay_classification,
  levels = c("on_time", "mild_delay", "moderate_delay", "severe_delay")
)

# Define color mapping similar to Python
color_map <- c(
  "on_time" = "green",       
  "mild_delay" = "yellow",    
  "moderate_delay" = "orange",
  "severe_delay" = "red"  
)

# Plot
ggplot(grouped_time_df, aes(x = time_of_day, y = percentage, fill = delay_classification)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8)) +
  geom_text(aes(label = paste0(percentage, "%")),
            position = position_dodge(width = 0.8),
            vjust = -0.3, size = 3.5) +
  scale_fill_manual(values = color_map) +
  labs(title = "Percentage of Delays by Time of Day",
       x = "Time of Day", y = "Percentage",
       fill = "Delay Classification") +
  theme_minimal()
```
Based on the graph above, morning flights have the most flights that are on time at 86.68% of flights, while night flights tend to be delayed more often with only 70.69% of flights being on time. Based on the visuals above, I would recommend booking tickets in the mornings and avoiding booking night flights.


# Finding the best day of the week to book flight tickets.

 In order to find the best day of the week to book a plane ticket, we will be grouping delays by "dayofweekstring" and 'delay_classification" and finding the count of each.
```{r}
# Grouping by day of week and delay classification, and counting
grouped_day_df <- final_delay_df %>%
  group_by(dayofweekstring, delay_classification) %>%
  summarise(count = n(), .groups = 'drop')

# Calculate percentages within each day of the week
grouped_day_df <- grouped_day_df %>%
  group_by(dayofweekstring) %>%
  mutate(percentage = round((count / sum(count)) * 100, 2)) %>%
  ungroup()

# View the result
print(grouped_day_df)
```
Plotting a bar chart that shows percentage of delays against day of the week.
```{r}
final_delay_df$dayofweekstring <- factor(
  final_delay_df$dayofweekstring,
  levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")
)

# Group, count, and calculate percentages
grouped_day_df <- final_delay_df %>%
  group_by(dayofweekstring, delay_classification) %>%
  summarise(count = n(), .groups = 'drop') %>%
  group_by(dayofweekstring) %>%
  mutate(percentage = round((count / sum(count)) * 100, 2)) %>%
  ungroup()

# Define color map
color_map <- c(
  "on_time" = "green",        # Green
  "mild_delay" = "yellow",     # Yellow
  "moderate_delay" = "orange", # Orange
  "severe_delay" = "red"    # Red
)

# Create bar plot
ggplot(grouped_day_df, aes(x = dayofweekstring, y = percentage, fill = delay_classification)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label = paste0(percentage, "%")), 
            position = position_dodge(width = 0.9), 
            vjust = -0.2, size = 3) +
  scale_fill_manual(values = color_map) +
  labs(title = "Percentage of Delays by Day of the Week",
       x = "Day of the Week",
       y = "Percentage",
       fill = "Delay Classification") +
  theme_minimal()

```
Based on the graph above, Saturdays has the highest percentage of flights that are on time at 82.98% while Fridays has the least percentage of flights that are on time at 76.12%. 

Based on the visual above, I would recommend booking flights on Saturdays while avoiding booking flights on Fridays.

# Finding best days of the week to book a plane ticket for each year
```{r}
# Group by year and day of week, and calculate average arrival delay
grouped_days <- final_delay_df %>%
  group_by(year, dayofweekstring) %>%
  summarise(avg_arrdelay = mean(arrdelay, na.rm = TRUE), .groups = "drop")

# Find the best day (with minimum average delay) for each year
best_days <- grouped_days %>%
  group_by(year) %>%
  filter(avg_arrdelay == min(avg_arrdelay)) %>%
  ungroup()

# Print results
for (i in 1:nrow(best_days)) {
  year <- best_days$year[i]
  day <- best_days$dayofweekstring[i]
  delay <- round(best_days$avg_arrdelay[i], 2)
  cat(sprintf("In %d, the best day to fly was %s with an average arrival delay of %.2f minutes.\n", year, day, delay))
}

```
# Finding best time of day to book a plane ticket for each year
```{r}
# Group by year and time of day, calculate average arrival delay
grouped_time <- final_delay_df %>%
  group_by(year, time_of_day) %>%
  summarise(avg_arrdelay = mean(arrdelay, na.rm = TRUE), .groups = "drop")

# For each year, find the time of day with the lowest average delay
best_times <- grouped_time %>%
  group_by(year) %>%
  filter(avg_arrdelay == min(avg_arrdelay)) %>%
  ungroup()

# Print results for each year
for (i in 1:nrow(best_times)) {
  year <- best_times$year[i]
  time <- best_times$time_of_day[i]
  delay <- round(best_times$avg_arrdelay[i], 2)
  cat(sprintf("In %d, the best time of day to fly was %s with an average arrival delay of %.2f minutes.\n", year, time, delay))
}

```
# Part 2(b) Evaluate wether older planes suffer more delays on a year-to-year basis.

Loading dataset
```{r}
planes_data_df <- read.csv("dataset/plane-data.csv")
print(planes_data_df)
```

Checking for null values and data type
```{r}
colSums(is.na(planes_data_df))
str(planes_data_df)
planes_data_df
```

Converting tailnums in plane_data_df to lowercase 
```{r}
planes_data_df$tailnum <- tolower(planes_data_df$tailnum)
```

Combining main_df and planes_data_df using 'tailnum'.
```{r}
merged_df <- main_df %>%
  left_join(planes_data_df, by = "tailnum", suffix = c("main_df", "planes_data_df"))

merged_df
```
Checking for null values in merged_df
```{r}
colSums(is.na(merged_df))
```

```{r}
merged_df
```
We will be choosing and renaming some columns to answer the second question.

Same as part (a), we will be using arrival delay only as DOT defines delay as arrival delay.
```{r}
age_df <- merged_df[, c("yearmain_df", "flightnum", "tailnum", "arrdelay", "yearplanes_data_df")]
head(age_df)
```
Rename yearmain_df to current_year and yearplanes_data_df to plane_year
```{r}
colnames(age_df)[colnames(age_df) == "yearmain_df"] <- "current_year"
colnames(age_df)[colnames(age_df) == "yearplanes_data_df"] <- "plane_year"

head(age_df)
```
Removing null values from plane_year
```{r}
age_df <- age_df[!is.na(age_df$plane_year), ]

colSums(is.na(age_df))
```
To find the age of a plane, we will be taking current year minus the issue date.
```{r}
#converting plane_year to integers before calculation
age_df$plane_year <- as.integer(age_df$plane_year)

#calculating plane age using current year minus issue date 
age_df$plane_age <- age_df$current_year - age_df$plane_year
head(age_df)


```

```{r}
colSums(is.na(age_df))
```

Removing null values in arrival delays and plane age
```{r}
age_df <- age_df[!is.na(age_df$arrdelay) & !is.na(age_df$plane_age), ]
colSums(is.na(age_df))

```
We will now be grouping the dataset by the plane's tail number to get the average delays for each model of plane.
```{r}
plane_summary_df <- age_df %>%
  group_by(tailnum) %>%
  summarise(
    average_arrdelay = mean(arrdelay, na.rm = TRUE),
    plane_age = first(plane_age)                       
  )
head(plane_summary_df)
```
Removing plane ages that are less than 0
```{r}
plane_summary_df <- plane_summary_df %>%
  filter(plane_age >= 0)

head(plane_summary_df)
```
We will now be grouping the dataset by the plane's age. 
```{r}
plane_age_groups <- plane_summary_df %>%
  group_by(plane_age) %>%
  summarise(average_arrdelay = mean(average_arrdelay, na.rm = TRUE))
head(plane_age_groups)
```
Plotting average arrival delay against plane age to see the general trend. 
```{r}
ggplot(plane_age_groups, aes(x = plane_age, y = average_arrdelay)) +
  geom_line(color = "red") +
  geom_point(color = "red") +
  labs(
    title = "Average Arrival Delay by Plane Age",
    x = "Plane Age",
    y = "Average Arrival Delay"
  ) +
  theme_minimal()
```
Based on the graph above, from age 0 to 24, the general trend of delays against plane age is decreasing, meaning as age increases, delay decreases. However, after the age of 24, delay is generally increasing as age increases.


```{r}
ggplot(plane_age_groups, aes(x = plane_age, y = average_arrdelay)) +
  geom_point() +  # Scatterplot points
  geom_smooth(method = "lm", color = "red") +  # Linear regression line
  labs(
    title = "Plane Age vs Average Arrival Delay with Trend Line",
    x = "Plane Age",
    y = "Average Arrival Delay"
  ) +
  theme_minimal()

```

```{r}
library(ggcorrplot)

# Compute Spearman correlation
corr_matrix <- cor(plane_age_groups, method = "spearman", use = "complete.obs")

# Plot
ggcorrplot(corr_matrix, lab = TRUE, lab_size = 4, 
           colors = c("blue", "white", "red"),
           title = "Feature Correlation Heatmap")

```
This shows that plane age has a negative correlation between arrival delays, which supports the graphs above, where as plane age increases, delays generally decrease. 


Reasons why this might be the case:

- newer planes being assigned to longer and more complicated routes while older planes being assigned to shorter and simpler routes, leading to newer planes having more delays compared to older planes

- older planes may be used much lesser than newer planes, leading to more instances of delays for newer planes

- there may be a survivorship bias as the older planes still flying are more reliable than the newer ones.


# Part 2(c) Logistic Regression for the probability of diverted US flights
Data Cleaning 
We will be first be removing cancelled flights as cancelled flights can't be diverted.

```{r}
diverted_df <- main_df
diverted_df <- diverted_df[diverted_df$cancelled == 0, ]
table(diverted_df$cancelled)
```
Removing null values in columns that we will be using for logistic regression.
```{r}
colSums(is.na(diverted_df))

```

```{r}
features <- c('month', 'dayofmonth', 'dayofweek', 'crsdeptime', 'crsarrtime', 'distance','uniquecarrier', 'origin', 'dest', 'depdelay', 'taxiout')

target <- 'diverted'

diverted_df <- diverted_df[!is.na(diverted_df[[target]]) & !apply(diverted_df[features], 1, anyNA), ]

colSums(is.na(diverted_df))
```
We will be first checking for perfect multicollinearity to identify and remove redundant variables.
```{r}
numerical_variables <- c('month', 'dayofmonth', 'dayofweek', 'crsdeptime', 'crsarrtime', 'distance', 'depdelay', 'taxiout')

colSums(is.na(diverted_df[numerical_variables]))
```

```{r}
library(ggcorrplot)

# Assign numerical variables to corr_df
corr_df <- diverted_df[, numerical_variables]

# Compute the Pearson correlation matrix
corr_matrix <- cor(corr_df, method = "pearson", use = "complete.obs")

# Plot the correlation heatmap
ggcorrplot(corr_matrix, lab = TRUE, lab_size = 4, 
           colors = c("blue", "white", "red"),
           title = "Pearson Correlation Heatmap (Numeric Features)")

```
Since there is no pairwise correlation that is equal to one, the assumption of no perfect multicollinerity is satisfied. However, we will be removing dayofmonth as it has very low correlation and has low predictive value.

# Logistic Regression Analysis throughout the years

For R, we will be splitting the dataset into its years so that there is enough memory to run the code.

```{r}
diverted_df
```
Selecting the columns 
```{r}
# Using features and target that are defined earlier
model_df <- diverted_df %>%
  select(all_of(features), all_of(target),year)
model_df
```


```{r}
print(colSums(is.na(model_df)))
```
Splitting the dataset by its year
```{r}
five_df <- model_df %>% filter(year == 1995)
six_df <- model_df %>% filter(year == 1996)
seven_df <- model_df %>% filter(year == 1997)
eight_df <- model_df %>% filter(year == 1998)
nine_df <- model_df %>% filter(year == 1999)

head(five_df)

```

Maximising the amount of memory
```{r}
# List of all objects currently in the environment
all_objects <- ls()

# List of objects to keep (for example, keep year-wise dataframes and other necessary elements for model)
items_to_keep <- c("five_df", "six_df", "seven_df", "eight_df", "nine_df")

# Identify objects to remove by subtracting the ones you want to keep
objects_to_remove <- setdiff(all_objects, items_to_keep)

# Remove the unnecessary objects from the environment
rm(list = objects_to_remove)

# Optionally, check that only the necessary objects remain
print(ls())

```

```{r}
# Load necessary libraries
library(dplyr)
library(caret)
library(ggplot2)
library(pROC)
library(ROCR)

# Define categorical and numerical features
cat_features <- c('month', 'dayofweek', 'uniquecarrier', 'origin', 'dest')
num_features <- c('crsdeptime', 'crsarrtime', 'distance', 'depdelay', 'taxiout')
target <- 'diverted'

# Prepare the dataset for the year 1995 (example for five_df)
df_year <- five_df %>%
  drop_na(all_of(c(cat_features, num_features, target)))

# Check if there's data after filtering
if (nrow(df_year) == 0) {
  stop("No data for year 1995")
}

# Split into X (features) and y (target)
X <- df_year[, c(cat_features, num_features)]
y <- df_year[, target]

# Train/test split with stratification
set.seed(42)
train_index <- createDataPartition(y, p = 0.7, list = FALSE)
X_train <- X[train_index, ]
y_train <- y[train_index]
X_test <- X[-train_index, ]
y_test <- y[-train_index]

# Preprocessing: Center and scale numeric features and apply one-hot encoding to categorical features
preProc <- preProcess(X_train, method = c("center", "scale"))
X_train <- predict(preProc, X_train)
X_test <- predict(preProc, X_test)

# Apply one-hot encoding for categorical features
X_train_cat <- model.matrix(~ month + dayofweek + uniquecarrier + origin + dest - 1, data = X_train)
X_test_cat <- model.matrix(~ month + dayofweek + uniquecarrier + origin + dest - 1, data = X_test)

# Combine the numerical and encoded categorical features
X_train_final <- cbind(X_train[, num_features], X_train_cat)
X_test_final <- cbind(X_test[, num_features], X_test_cat)

# Logistic Regression Model
model <- glm(y_train ~ ., data = as.data.frame(X_train_final), family = binomial(link = 'logit'))

# Predictions and probabilities
y_pred <- predict(model, as.data.frame(X_test_final), type = "response")
y_pred_class <- ifelse(y_pred > 0.5, 1, 0)

# Evaluate the model using confusion matrix, accuracy, and F1 score
confusion <- confusionMatrix(as.factor(y_pred_class), as.factor(y_test))

print("Year 1995 Results")
print(confusion)

# ROC Curve
roc_obj <- roc(y_test, y_pred)
plot(roc_obj, col = "blue", main = "ROC Curve for Year 1995")
print(paste("AUC:", auc(roc_obj)))


```



```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```









































We will first be splitting our features into categorical and numerical features
```{r}
# Categorical features
cat_features <- c('month', 'dayofweek', 'uniquecarrier', 'origin', 'dest')

# Numerical features
num_features <- c('crsdeptime', 'crsarrtime', 'distance', 'depdelay', 'taxiout')
```

Using a for loop, cycle throughout the years and apply logistic regeression to it
```{r}
library(dplyr)
library(caret)
library(ggplot2)
library(pROC)
library(ROCR)
library(tidyr)

# Define features and target
features <- c('month', 'dayofweek', 'uniquecarrier', 'origin', 'dest', 'crsdeptime', 'crsarrtime', 'distance', 'depdelay', 'taxiout')
target <- 'diverted'

# Function to preprocess and fit logistic regression model
process_and_fit_model <- function(df_year) {
  # Preprocessing: Center and scale numeric features
  preProc <- preProcess(df_year[, features], method = c("center", "scale"))
  df_year_scaled <- predict(preProc, df_year[, features])

  # Logistic Regression Model
  model <- glm(diverted ~ ., data = cbind(df_year_scaled, diverted = df_year[, target]), family = binomial(link = 'logit'))
  
  # Store coefficients
  coef_values <- coef(model)
  coef_df <- data.frame(feature = names(coef_values), coefficient = coef_values, year = df_year$year[1])

  # Predictions
  y_pred <- predict(model, df_year_scaled, type = "response")
  y_pred_class <- ifelse(y_pred > 0.5, 1, 0)
  
  # Confusion Matrix and ROC Curve
  confusion <- confusionMatrix(as.factor(y_pred_class), as.factor(df_year[, target]))
  roc_obj <- roc(df_year[, target], y_pred)
  
  list(coef_df = coef_df, confusion = confusion, roc_obj = roc_obj)
}

# Empty list to store coefficients for each year
coef_list <- list()

# Loop through years to train and evaluate the model for each year
for (year in sort(unique(diverted_df$year))) {
  
  # Filter data for current year and drop rows with NA in selected features and target
  df_year <- diverted_df %>%
    filter(year == year) %>%
    drop_na(all_of(c(features, target)))
  
  if (nrow(df_year) > 0) {
    result <- process_and_fit_model(df_year)
    
    # Store coefficients
    coef_list[[length(coef_list) + 1]] <- result$coef_df
    
    # Print confusion and ROC AUC for this year
    print(paste("\n--- Year", year, "---"))
    print(result$confusion)
    plot(result$roc_obj, col = "blue", main = paste("ROC Curve - Year", year))
    print(paste("AUC:", auc(result$roc_obj)))
  }
}

# Combine all coefficients into one data frame
coef_df_final <- do.call(rbind, coef_list)

# Plot the coefficients or save them
if (nrow(coef_df_final) > 0) {
  ggplot(coef_df_final, aes(x = feature, y = coefficient, color = factor(year))) + 
    geom_bar(stat = "identity") +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
    labs(title = "Logistic Regression Coefficients per Year", x = "Features", y = "Coefficient")
} else {
  print("No coefficients to plot.")
}



```

```{r}
library(ggplot2)
library(dplyr)

# Combine all coefficients into one data frame
coeffs_df <- do.call(rbind, coeff)

# Choose key features to visualize
features_to_plot <- c('num__distance', 'cat__month_2', 'cat__dayofweek_2', 'num__depdelay', 'num__taxiout')

# Filter the coefficients for the features of interest
subset_coeffs <- coeffs_df %>% filter(feature %in% features_to_plot)

# Plot the coefficients over time
ggplot(subset_coeffs, aes(x = year, y = coefficient, color = feature, group = feature)) +
  geom_line() +
  geom_point() +  # Add markers
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray", size = 1) +  # Reference line at y = 0
  labs(
    title = "Logistic Regression Coefficients Over Time",
    x = "Year",
    y = "Coefficient"
  ) +
  theme_minimal() +
  theme(legend.position = "right") +
  guides(color = guide_legend(title = "Feature"))

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```